---
title: "PHP_1511_Final"
author: "Linda Wakamoto"
date: "2024-08-14"
output: html_document
---
```{r}
title: "PHP1511 Final Project"
author: "Linda Wakamoto"
date: '2023-04-14'
output: html_document
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nhanesA) #DEMO = demographics, DIET = dietary, EXAM = examination, LAB = laboratory, Q = questionnaire
nhanesTables(data_group='DEMO',year=2015)
demo <- nhanes('DEMO_I')
names(demo)
table(demo$DMDHSEDU)
cumsum(table(demo$DMDHSEDU))
length(is.na(demo$DMDHSEDU))
library(dplyr)

#Age
age_1999 <- nhanes('DEMO')[c('SEQN','RIDAGEYR')]
age_2001 <- nhanes('DEMO_B')[c('SEQN','RIDAGEYR')]
age_2003 <- nhanes('DEMO_C')[c('SEQN','RIDAGEYR')]
age_2005 <- nhanes('DEMO_D')[c('SEQN','RIDAGEYR')]
age_2007 <- nhanes('DEMO_E')[c('SEQN','RIDAGEYR')]
age_2009 <- nhanes('DEMO_F')[c('SEQN','RIDAGEYR')]
age_2011 <- nhanes('DEMO_G')[c('SEQN','RIDAGEYR')]
age_2013 <- nhanes('DEMO_H')[c('SEQN','RIDAGEYR')]
age_2015 <- nhanes('DEMO_I')[c('SEQN','RIDAGEYR')]
age_2017 <- nhanes('DEMO_J')[c('SEQN','RIDAGEYR')]

all_age <- bind_rows(age_1999, age_2001, age_2003, age_2005, age_2007, age_2009, age_2011, age_2013, age_2015, age_2017)
colnames(all_age) <- c("PATIENT_ID","Age")
```

```{r}
#male
#1 for male, 2 for female
gender_1999 <- nhanes('DEMO')[c('SEQN','RIAGENDR')]
gender_2001 <- nhanes('DEMO_B')[c('SEQN','RIAGENDR')]
gender_2003 <- nhanes('DEMO_C')[c('SEQN','RIAGENDR')]
gender_2005 <- nhanes('DEMO_D')[c('SEQN','RIAGENDR')]
gender_2007 <- nhanes('DEMO_E')[c('SEQN','RIAGENDR')]
gender_2009 <- nhanes('DEMO_F')[c('SEQN','RIAGENDR')]
gender_2011 <- nhanes('DEMO_G')[c('SEQN','RIAGENDR')]
gender_2013 <- nhanes('DEMO_H')[c('SEQN','RIAGENDR')]
gender_2015 <- nhanes('DEMO_I')[c('SEQN','RIAGENDR')]
gender_2017 <- nhanes('DEMO_J')[c('SEQN','RIAGENDR')]

all_gender <- bind_rows(gender_1999, gender_2001, gender_2003, gender_2005, gender_2007, gender_2009, gender_2011, gender_2013, gender_2015, gender_2017)
colnames(all_gender) <- c("PATIENT_ID","Male")
```

```{r}
#race
#1 for Mexican American
#2 for Other Hispanic
#3 for Non-hispanic white
#4 for Non-hispanic black
#5 for Other Race - Including Multi-Racial
race_1999 <- nhanes('DEMO')[c('SEQN','RIDRETH1')]
race_2001 <- nhanes('DEMO_B')[c('SEQN','RIDRETH1')]
race_2003 <- nhanes('DEMO_C')[c('SEQN','RIDRETH1')]
race_2005 <- nhanes('DEMO_D')[c('SEQN','RIDRETH1')]
race_2007 <- nhanes('DEMO_E')[c('SEQN','RIDRETH1')]
race_2009 <- nhanes('DEMO_F')[c('SEQN','RIDRETH1')]
race_2011 <- nhanes('DEMO_G')[c('SEQN','RIDRETH1')]
race_2013 <- nhanes('DEMO_H')[c('SEQN','RIDRETH1')]
race_2015 <- nhanes('DEMO_I')[c('SEQN','RIDRETH1')]
race_2017 <- nhanes('DEMO_J')[c('SEQN','RIDRETH1')]

all_race <- bind_rows(race_1999, race_2001, race_2003, race_2005, race_2007, race_2009, race_2011, race_2013, race_2015, race_2017)
colnames(all_race) <- c("PATIENT_ID","Race")
```

```{r}
#Education Level
edu_1999 <- nhanes('DEMO')[c('SEQN','DMDEDUC2')]
edu_2001 <- nhanes('DEMO_B')[c('SEQN','DMDEDUC2')]
edu_2003 <- nhanes('DEMO_C')[c('SEQN','DMDEDUC2')]
edu_2005 <- nhanes('DEMO_D')[c('SEQN','DMDEDUC2')]
edu_2007 <- nhanes('DEMO_E')[c('SEQN','DMDEDUC2')]
edu_2009 <- nhanes('DEMO_F')[c('SEQN','DMDEDUC2')]
edu_2011 <- nhanes('DEMO_G')[c('SEQN','DMDEDUC2')]
edu_2013 <- nhanes('DEMO_H')[c('SEQN','DMDEDUC2')]
edu_2015 <- nhanes('DEMO_I')[c('SEQN','DMDEDUC2')]
edu_2017 <- nhanes('DEMO_J')[c('SEQN','DMDEDUC2')]

all_edu <- bind_rows(edu_1999, edu_2001, edu_2003, edu_2005, edu_2007, edu_2009, edu_2011, edu_2013, edu_2015, edu_2017)
colnames(all_edu) <- c("PATIENT_ID","Education level")
```

```{r}
#ratio of family income to poverty
ratio_1999 <- nhanes('DEMO')[c('SEQN','INDFMPIR')]
ratio_2001 <- nhanes('DEMO_B')[c('SEQN','INDFMPIR')]
ratio_2003 <- nhanes('DEMO_C')[c('SEQN','INDFMPIR')]
ratio_2005 <- nhanes('DEMO_D')[c('SEQN','INDFMPIR')]
ratio_2007 <- nhanes('DEMO_E')[c('SEQN','INDFMPIR')]
ratio_2009 <- nhanes('DEMO_F')[c('SEQN','INDFMPIR')]
ratio_2011 <- nhanes('DEMO_G')[c('SEQN','INDFMPIR')]
ratio_2013 <- nhanes('DEMO_H')[c('SEQN','INDFMPIR')]
ratio_2015 <- nhanes('DEMO_I')[c('SEQN','INDFMPIR')]
ratio_2017 <- nhanes('DEMO_J')[c('SEQN','INDFMPIR')]

all_ratio <- bind_rows(ratio_1999, ratio_2001, ratio_2003, ratio_2005, ratio_2007, ratio_2009, ratio_2011, ratio_2013, ratio_2015, ratio_2017)
colnames(all_ratio) <- c("PATIENT_ID","Ratio of family income to poverty")
```

```{r}
#alcohol use
alc_1999 <- nhanes('ALQ')[c('SEQN','ALQ110')]
alc_2001 <- nhanes('ALQ_B')[c('SEQN','ALQ110')]
alc_2003 <- nhanes('ALQ_C')[c('SEQN','ALQ110')]
alc_2005 <- nhanes('ALQ_D')[c('SEQN','ALQ110')]
alc_2007 <- nhanes('ALQ_E')[c('SEQN','ALQ110')]
alc_2009 <- nhanes('ALQ_F')[c('SEQN','ALQ110')]
alc_2011 <- nhanes('ALQ_G')[c('SEQN','ALQ110')]
alc_2013 <- nhanes('ALQ_H')[c('SEQN','ALQ110')]
alc_2015 <- nhanes('ALQ_I')[c('SEQN','ALQ110')]
alc_2017 <- nhanes('ALQ_J')[c('SEQN','ALQ111')] 

all_alc <- bind_rows(alc_1999, alc_2001, alc_2003, alc_2005, alc_2007, alc_2009, alc_2011, alc_2013, alc_2015, alc_2017)
colnames(all_alc) <- c("PATIENT_ID","Alcohol use")
```

```{r}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(tableone)
library(gtsummary)
```

Upload data
```{r}
data <- read.csv('nhanes_data.csv')

# remove missing data from lead or hypertension
data <- data %>% drop_na(Lead..ug.dl.)
data <- data %>% drop_na(Hypertension)

# 18+ age only
data <- subset(data, Age >=20 )

# omit education, ratio, BMI, (alc, smoking)
data <- data %>% drop_na(all_edu)
data <- data %>% drop_na(BMI..kg.m.2.)
data <- data %>% drop_na(Ratio.of.family.income.to.poverty) # 38424 data existing...too much missing data with alc and smoking

# make categorical variables factors
factors <- c('Alcohol.use', 'all_edu', 'Male', 'Race', 'all_smoking', 'Hypertension')
data[,factors] <- lapply(data[,factors], factor)
```

Make stratified tables (Baseline)
```{r}
summary(data)
# q1: <0.830; q2: 0.830 - 1.772; q3: 1.772-2.140; q4: 2.140+
quant <- cut(data$Lead..ug.dl., quantile(data$Lead..ug.dl., probs=0:4/4), include.lowest=TRUE)
data <- data.frame(data, quant)
data$quant <- as.factor(data$quant)
baseline_table1 <- tbl_summary(data)
baseline_table2 <- tbl_summary(data=data, by='quant') %>%
  add_p()


merged <- tbl_merge(tbls=list(baseline_table1, baseline_table2),
                    tab_spanner = c("**All Data**", "**Quantiles**"))
merged
```

Develop logistic regression models - lead
```{r}
lead1 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit')) %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))

lead2 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race, family=binomial(link='logit')) %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))

lead3 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit')) %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))

lead_mod <- tbl_merge(tbls=list(lead1, lead2, lead3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))
lead_mod
```
quartiles
```{r}
quant11 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=quant=='[0.05,0.83]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant21 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race, family=binomial(link='logit'), subset=quant=='[0.05,0.83]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant31 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=quant=='[0.05,0.83]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant_mod1 <- tbl_merge(tbls=list(quant11, quant21, quant31), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))


quant12 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=quant=='(0.83,1.33]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant22 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race, family=binomial(link='logit'), subset=quant=='(0.83,1.33]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant32 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=quant=='(0.83,1.33]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant_mod2 <- tbl_merge(tbls=list(quant12, quant22, quant32), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))


quant13 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=quant=='(1.33,2.14]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant23 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race, family=binomial(link='logit'), subset=quant=='(1.33,2.14]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant33 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=quant=='(1.33,2.14]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant_mod3 <- tbl_merge(tbls=list(quant13, quant23, quant33), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))


quant14 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=quant=='(2.14,61.3]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant24 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race, family=binomial(link='logit'), subset=quant=='(2.14,61.3]') %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
quant34 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=quant=='(2.14,61.3]') %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
quant_mod4 <- tbl_merge(tbls=list(quant14, quant24, quant34), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

```
gender
```{r}
male11 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=Male==1) %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
male21 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Race, family=binomial(link='logit'), subset=Male==1) %>%
  tbl_regression(exponentiate=TRUE, include=c(Lead..ug.dl.))
male31 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Race+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=Male==1) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
male_mod1 <- tbl_merge(tbls=list(male11, male21, male31), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))


male12 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=Male==2) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
male22 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Race, family=binomial(link='logit'), subset=Male==2) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
male32 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Race+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=Male==2) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
male_mod2 <- tbl_merge(tbls=list(male12, male22, male32), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

```

```{r}
race11 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=Race==1) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race21 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male, family=binomial(link='logit'), subset=Race==1) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race31 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=Race==1) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.),)
race_mod1 <- tbl_merge(tbls=list(race11, race21, race31), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))


race12 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=Race==2) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race22 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male, family=binomial(link='logit'), subset=Race==2) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race32 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=Race==2) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race_mod2 <- tbl_merge(tbls=list(race12, race22, race32), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))


race13 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=Race==3) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race23 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male, family=binomial(link='logit'), subset=Race==3) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race33 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=Race==3) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race_mod3 <- tbl_merge(tbls=list(race13, race23, race33), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))


race14 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=Race==4) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race24 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male, family=binomial(link='logit'), subset=Race==4) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race34 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=Race==4) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race_mod4 <- tbl_merge(tbls=list(race14, race24, race34), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))


race15 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=Race==5) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race25 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male, family=binomial(link='logit'), subset=Race==5) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race35 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=Race==5) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
race_mod5 <- tbl_merge(tbls=list(race15, race25, race35), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))
```

```{r}
bmi11 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=BMI..kg.m.2.>=30) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
bmi21 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race, family=binomial(link='logit'), subset=BMI..kg.m.2.>=30) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
bmi31 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race+all_edu+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=BMI..kg.m.2.>=30) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
bmi_mod1 <- tbl_merge(tbls=list(bmi11, bmi21, bmi31), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))


bmi12 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=(BMI..kg.m.2.>25 & BMI..kg.m.2.<=30)) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
bmi22 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race, family=binomial(link='logit'), subset=(BMI..kg.m.2.>25 & BMI..kg.m.2.<=30)) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
bmi32 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race+all_edu+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=(BMI..kg.m.2.>25 & BMI..kg.m.2.<=30)) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
bmi_mod2 <- tbl_merge(tbls=list(bmi12, bmi22, bmi32), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))


bmi13 <- glm(data=data, formula=Hypertension~Lead..ug.dl., family=binomial(link='logit'), subset=BMI..kg.m.2.<=25) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
bmi23 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race, family=binomial(link='logit'), subset=BMI..kg.m.2.<=25) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
bmi33 <- glm(data=data, formula=Hypertension~Lead..ug.dl.+Age+Male+Race+all_edu+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, family=binomial(link='logit'), subset=BMI..kg.m.2.<=25) %>%
  tbl_regression(exponentiate=TRUE,include=c(Lead..ug.dl.))
bmi_mod3 <- tbl_merge(tbls=list(bmi13, bmi23, bmi33), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))
```

```{r}
final_summary <- tbl_stack(tbls=list(lead_mod, quant_mod1, quant_mod2, quant_mod3, quant_mod4, male_mod1, male_mod2, race_mod1, race_mod2, race_mod3, race_mod4, race_mod5, bmi_mod1, bmi_mod2, bmi_mod3))
final_summary
```
## Making Table 3

Per BLL increase
```{r}
lead1_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.) %>%
  tbl_regression(include=c(Lead..ug.dl.))

lead2_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+Male+Race) %>%
  tbl_regression(include=c(Lead..ug.dl.))

lead3_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+Race+Male+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use) %>%
  tbl_regression(include=c(Lead..ug.dl.))

lead_mod_t3 <- tbl_merge(tbls=list(lead1_t3, lead2_t3, lead3_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))
lead_mod_t3
```

Stratified by race and sex
```{r}
#Race 1 Male 1
m_race11_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl., subset=Race==1 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.))
m_race21_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age, subset=Race==1 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.))
m_race31_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, subset=Race==1 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.),)
m_race_mod1_t3 <- tbl_merge(tbls=list(m_race11_t3, m_race21_t3, m_race31_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

#Race 1 Male 2
f_race11_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl., subset=Race==1 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.))
f_race21_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age, subset=Race==1 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.))
f_race31_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, subset=Race==1 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.),)
f_race_mod1_t3 <- tbl_merge(tbls=list(f_race11_t3, f_race21_t3, f_race31_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

#Race 2 Male 1
m_race12_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl., subset=Race==2 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.))
m_race22_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age, subset=Race==2 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.))
m_race32_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, subset=Race==2 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.),)
m_race_mod2_t3 <- tbl_merge(tbls=list(m_race12_t3, m_race22_t3, m_race32_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

#Race 2 Male 2
f_race12_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl., subset=Race==2 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.))
f_race22_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age, subset=Race==2 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.))
f_race32_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, subset=Race==2 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.),)
f_race_mod2_t3 <- tbl_merge(tbls=list(f_race12_t3, f_race22_t3, f_race32_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

#Race 3 Male 1
m_race13_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl., subset=Race==3 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.))
m_race23_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age, subset=Race==3 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.))
m_race33_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, subset=Race==3 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.),)
m_race_mod3_t3 <- tbl_merge(tbls=list(m_race13_t3, m_race23_t3, m_race33_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

#Race 3 Male 2
f_race13_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl., subset=Race==3 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.))
f_race23_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age, subset=Race==3 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.))
f_race33_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, subset=Race==3 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.),)
f_race_mod3_t3 <- tbl_merge(tbls=list(f_race13_t3, f_race23_t3, f_race33_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

#Race 4 Male 1
m_race14_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl., subset=Race==4 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.))
m_race24_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age, subset=Race==4 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.))
m_race34_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, subset=Race==4 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.),)
m_race_mod4_t3 <- tbl_merge(tbls=list(m_race14_t3, m_race24_t3, m_race34_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

#Race 4 Male 2
f_race14_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl., subset=Race==4 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.))
f_race24_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age, subset=Race==4 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.))
f_race34_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, subset=Race==4 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.),)
f_race_mod4_t3 <- tbl_merge(tbls=list(f_race14_t3, f_race24_t3, f_race34_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

#Race 5 Male 1
m_race15_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl., subset=Race==5 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.))
m_race25_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age, subset=Race==5 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.))
m_race35_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, subset=Race==5 & Male==1) %>%
  tbl_regression(include=c(Lead..ug.dl.),)
m_race_mod5_t3 <- tbl_merge(tbls=list(m_race15_t3, m_race25_t3, m_race35_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))

#Race 5 Male 2
f_race15_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl., subset=Race==5 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.))
f_race25_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age, subset=Race==5 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.))
f_race35_t3 <- lm(data=data, formula=avg_sys~Lead..ug.dl.+Age+all_edu+BMI..kg.m.2.+Ratio.of.family.income.to.poverty+all_smoking+Alcohol.use, subset=Race==5 & Male==2) %>%
  tbl_regression(include=c(Lead..ug.dl.),)
f_race_mod5_t3 <- tbl_merge(tbls=list(f_race15_t3, f_race25_t3, f_race35_t3), tab_spanner=c('Model 1', 'Model 2', 'Model 3'))
```

```{r}
final_summary_t3 <- tbl_stack(tbls=list(lead_mod_t3, m_race_mod1_t3, m_race_mod2_t3, m_race_mod3_t3, m_race_mod4_t3, m_race_mod5_t3, f_race_mod1_t3, f_race_mod2_t3, f_race_mod3_t3, f_race_mod4_t3, f_race_mod5_t3))
final_summary_t3
```

